---
title: "p8130_hw2_yl5508"
author: "Yifei LIU (yl5508)"
date: 2023/10/11
output: 
  html_document
---

```{r setting, echo = FALSE}
options(scipen = 999)
```

## Problem 1

**(a)** $P(exactly\ 40)=\binom{56}{40} \cdot (0.73)^{40} \cdot (1 - 0.73)^{16}=0.113=11.3\%$  

**(b)** 
```{r p1(b)}
trial = 56
success = 39
p = 0.73
print(1-pbinom(success,trial,p))
```
The probability that at least 40 of them have at least 1 checkup is `r 1-pbinom(success,trial,p)`.  

**(c)** 
**(1)** The condition we should check before using Poisson appoximation to binomial is:  
- n must be large (n>100): in this problem, n=56<100.  
- Probability of success p should be small (p<0.01): in this problem, p is 0.73.  
As a result, we should not use Poisson as an approximation to binomial.  
**(2)** Try to use normal approximation:  
- np≥10: $np=56\cdot0.73=40.88>10$  
- n(1-p)≥10: $n(1-p)=56\cdot(1-0.73)=15.12>10$  
So, we can use normal approximation to binomial. The normal approximation distribution would be $N~(40.88,11.04)$  
```{r p1(c)}
mu_norm = 56*0.73
sd_norm = sqrt(56*0.73*(1-0.73))
norm_eql40 = pnorm(40.5,mu_norm,sd_norm)-pnorm(39.5,mu_norm,sd_norm)
norm_leq40 = 1-pnorm(39.5,mu_norm,sd_norm)
print(c(mu_norm,sd_norm,norm_eql40,norm_leq40))
```
For problem (a): using normal approximation to binomial, we get possibility 11.6%.  
For problem (b): using normal approximation to binomial, we get possibility 66.1%.  
The results shown above is quite close to the results we have calculated from binomial.  

**(d)** Expectation of a binomial R.V. is `E[X] = np`, which is `r trial*p`.  

**(e)** Variance of a binomial R.V. is `Var(x) = np(1-p)`, which is `r trial*p*(1-p)`. The standard deviation of it would be `r sqrt(trial*p*(1-p))`($sd=\sqrt{Var(x)}$).  

## Problem 2

**(a)**
```{r p2(a)}
print(ppois(2,6))
```
The PMF of Poisson distribution can be expressed as $P(X = k) = \frac{e^{-\lambda} \cdot \lambda^k}{k!}$.  
For tornado happened fewer than 3 times, $P(X < 3) = P(X = 0) + P(X = 1) + P(X = 2) = \sum_{k=0}^{2}\frac{e^{-6} \cdot 6^k}{k!} = 0.062 = 6.2\%$.  

**(b)** 
```{r p2(b)}
print(dpois(3,6))
```
The probability that tornado will happen exactly 3 tims is `r dpois(3,6)`.  

**(c)** 
```{r p2(c)}
print(1-ppois(3,6))
```
The probability that tornado will happen more than 3 tims is `r 1-ppois(3,6)`.  

## Problem 3

**(a)** 
```{r p3(a)}
print(1-pnorm(137,128,10.2))
```
The probability of a selected American man (20-29) with systolic blood pressure above 137.0 is `r 1-pnorm(137,128,10.2)`.  

**(b)**
```{r p3(b)}
print(pnorm(125,128,10.2/sqrt(50)))
```
The population distribution of the R.V. is normal and the sample size is larger than 30. So, we can denote the sampling distribution as $\overline{X}\sim N(\mu,\frac{\sigma^{2}}{n})$, which is $\overline{X}\sim N(128,\frac{10.2}{\sqrt{50}})$ in this problem. The probability that sample mean will be less than 125.0 is `r pnorm(125,128,10.2/sqrt(50))`.  

**(c)** 
```{r p3(c)}
print(qnorm(0.9,128,10.2/sqrt(40)))
```
For sample size of 40, the sampling distribution would follow: $\overline{X}\sim N(128,\frac{10.2}{\sqrt{40}})$. The 90th percentile of the sampling distribution is `r qnorm(0.9,128,10.2/sqrt(40))`.  

## Problem 4

**(a)** 
```{r p4(a)}
#t statistic
t_975 = qt(0.975,39)
print(t_975)
```
For true σ from population is unknown, we need to calculate the estimated standard error: $\frac{s}{\sqrt{n}}=\frac{10}{\sqrt{40}}=1.58$.  
So, the confidence interval(CI) for the population mean pulse rate would be shown as $P(\overline{X}-2.02\sigma_{\overline{X}}<\mu<\overline{X}+2.02\sigma_{\overline{X}}) = 0.95$, among which CI is **$76.81<\mu<83.19$**.  

**(b)** The confidence interval calculated above means that, for most (95%) of the random variables, $\overline{X}$ will fall within +/- 2.02 estimated SE of the true mean $\mu$. The 95% CI is $(76.81,83.19)$.  

**(c)** 
```{r p4(c)}
#statistic t
t = abs((80-70)/(10/sqrt(40)))
t_995 = qt(0.995,39)
print(c(t,t_995))
```
Using method of tests for the mean of a normal distribution with known variance.  
Suppose that $\mu_0=70$, and $H_0:\mu=\mu_0$ vs $H_1:\mu\neq\mu_0$.  
With the significant level $\alpha=0.01$, test statistic is $t=\frac{\overline{X}-\mu_0}{s/\sqrt{n}}=\frac{80-70}{10/\sqrt{40}}=6.32$.  
Criteria:  
- Reject $H_0$: if $|t|>t_{39,0.995}$,  
- Fail to reject $H_0$: if $|t|\leq t_{39,0.995}$. 
$t_{39,0.995}=2.71$.  
Cause $|t|>t_{39,0.995}$, we can reject $H_0$.  
Result: with significant level $\alpha=0.01$, we should reject $H_0$ and the mean pulse of young women suffering from fibromyalagia isnot equal to 70. In other words, the possibility of getting a sample with a mean equivalent to 80 or more extreme than that is less than 1%, given that the population mean is 70.  

