---
title: "p8130_hw2_yl5508"
author: "Yifei LIU (yl5508)"
date: 2023/10/11
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setting, echo = FALSE}
options(scipen = 999)
options(digits = 2)
```


## Problem 1

**(a)** $P(exactly\ 40)=\binom{56}{40} \cdot (0.73)^{40} \cdot (1 - 0.73)^{16}=0.113=11.3\%$  
**(b)** 
```{r p1(b)}
trial = 56
success = 39
p = 0.73
print(1-pbinom(success,trial,p))
```
The probability that at least 40 of them have at least 1 checkup is `r 1-pbinom(success,trial,p)`.  
**(c)** The condition we should check before using Poisson appoximation to binomial is:  
- n must be large (n>100): in this problem, n=56<100.  
- Probability of success p should be small (p<0.01): in this problem, p is 0.73.  
As a result, we should not use Poisson as an approximation to binomial.  
**(d)** Expectation of a binomial R.V. is `E[X] = np`, which is `r trial*p`.  
**(e)** Variance of a binomial R.V. is `Var(x) = np(1-p)`, which is `r trial*p*(1-p)`. The standard deviation of it would be `r sqrt(trial*p*(1-p))`($sd=\sqrt{}$).  

## Problem 2

**(a)**
```{r p2(a)}
print(ppois(2,6))
```
The PMF of Poisson distribution can be expressed as $P(X = k) = \frac{e^{-\lambda} \cdot \lambda^k}{k!}$.  
For tornado happened fewer than 3 times, $P(X < 3) = P(X = 0) + P(X = 1) + P(X = 2) = \sum_{k=0}^{2}\frac{e^{-6} \cdot 6^k}{k!} = 0.062 = 6.2\%$.  
**(b)** 
```{r p2(b)}
print(dpois(3,6))
```
The probability that tornado will happen exactly 3 tims is `r dpois(3,6)`.  
**(c)** 
```{r p2(c)}
print(1-ppois(3,6))
```
The probability that tornado will happen more than 3 tims is `r 1-ppois(3,6)`.  

## Problem 3

**(a)** 
```{r p3(a)}
print(1-pnorm(137,128,10.2))
```
The probability of a selected American man (20-29) with systolic blood pressure above 137.0 is `r 1-pnorm(137,128,10.2)`.  
**(b)**
```{r p3(b)}
print(pnorm(125,128,10.2/sqrt(50)))
```
The population distribution of the R.V. is normal and the sample size is larger than 30. So, we can denote the sampling distribution as $\overline{X}\sim N(\mu,\frac{\sigma^{2}}{n})$, which is $\overline{X}\sim N(128,\frac{10.2}{\sqrt{50}})$ in this problem. The probability that sample mean will be less than 125.0 is `r pnorm(125,128,10.2/sqrt(50))`.  
**(c)** 
```{r p3(c)}
print(qnorm(0.9,128,10.2/sqrt(40)))
```
For sample size of 40, the sampling distribution would follow: $\overline{X}\sim N(128,\frac{10.2}{\sqrt{40}})$. The 90th percentile of the sampling distribution is `r qnorm(0.9,128,10.2/sqrt(40))`.  

## Problem 4

**(a)** For sample size of 40 (>30), the sampling distribution would follow: $\overline{X}\sim N(80,10)$. In this case, standard error(SE) is $\sigma_{\overline{X}}=\frac{\sigma^{2}}{n}=10$.  
So, the confidence interval(CI) for the population mean pulse rate would be shown as $P(\mu-2\sigma_{\overline{X}}<X<\mu+2\sigma_{\overline{X}}) = 0.95$, among which CI is $60<X<100$.  
**(b)** The confidence interval calculated above means that, for most (95%) of the random variables, $\overline{X}$ will fall within +/- 2SE of the true mean $\mu$.  
**(c)** Using method of tests for the mean of a normal distribution with known variance.  
Suppose that $\mu_0=70$, and $H_0:\mu=\mu_0$ vs $H_1:\mu\neq\mu_0$.  
With the significant level $\alpha=0.01$, test statistic is $z=\frac{\overline{X}-\mu_0}{\sigma/\sqrt{n}}=\frac{80-70}{10}=1$.  
Criteria:  
- Reject $H_0$: if $|z|>z_{1-0.05}$,  
- Fail to reject $H_0$: if $|z|\leq z_{1-0.05}$. 
```{r p4(c)}
print(qnorm(0.95,0,1))
```
$z_{1-0.05} = z_{0.95}=1.64$.  
Cause $z<z_{1-0.05}$, we fail to reject $H_0$.  
With significant level $\alpha=0.01$, we cannot reject $H_0$ and the mean pulse of young women suffering from fibromyalagia is equal to 70.

